---
title: |
  Adopting open source practices  
  for better science
author: Pierce Edmiston
output:
  beamer_presentation:
    theme: metropolis
---

# Outline

Open source practices that make for more reproducible science:

1. Version control
2. Dynamic documents
3. Building from source

Conclusion: It's worth it!

```{r, include=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE)
source("presentation.R")
```

# Why I care about reproducibility

1. I want my research to be reproducible.
2. I want to attract collaborators.

# Steps toward reproducible science

- Blind data analysts to experimental conditions.
- Improve statistics education (adapted for web).
- Hire methodological consultants.
- Seek collaboration for scalability.

# Why I think open source is the answer

Compare these two goals of reproducibility in science and in open source:

1. Fellow researchers should be able to reproduce my work.  
2. Anyone should be able to use and contribute to this project.

# Version control

```{r version-control-in-the-wild, fig.width=12}
grid.arrange(img("undo-button"), img("wiki-revisions"), img("git-log"), nrow = 1)
```

# Pick your poison

- **git**
- mercurial
- subversion
- gitless

# Tools for climbing

```{r climbing-tools}
img("climbing-tools", ext = ".jpg", draw = TRUE)
```

# Conquer clutter

```{r cluttered-folder}
img("cluttered-folder", draw = TRUE)
```

# Version control's dirty little secret

(It only really works on plaintext files.)

```{r excel-panic, fig.cap="Excel panic. Well, did you make changes or didn't you??"}
img("excel-panic", draw = TRUE)
```

# The good news

Once you're working in plaintext, you can do lots of cool things.

- Full power of VCS (merge, blame, etc).
- Use free and open source tools (Unix).
- **Write dynamic documents.**


# Dynamic documents

- Philosophy: DRY, Literate Programming
- Tools: Sphinx, Jupyter, Knitr, Pandoc


# Philosophy

Don't Repeat Yourself (DRY)
: Every piece of knowledge must have a single, unambiguous, authoritative 
  representation within a system.

Literate programming (LP)
: Intermingle prose and code for better understanding of the program. The 
  explanation of a program does not need to resemble the program structure.

# Knitr

**Elegant, flexible and fast dynamic report generation with R.**

    Participants in condition A outperformed participants in
    condition B, `report_model_results(mod, param = "condition")`.

```{r knitr-side-by-side}
grid.arrange(img("knitr-source"), img("knitr-output"), nrow = 1)
```

# Dynamic documents in practice

- Handouts
- Homework
- Supplemental materials
- Conference proceedings
- Journal papers

# A litmus test for reproducible research

Can you build the published paper without the original data?
